---
title: "Exercise 3"
author: "Soo Jee Choi, Annie Nguyen, and Tarini Sudhakar"
date: "`r Sys.Date()`"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)
```

##  1. Clustering and PCA

##  2. Market segmentation
Your task to is analyze this data as you see fit, and to prepare a (short!) report for NutrientH20 that identifies any interesting market segments that appear to stand out in their social-media audience. You have complete freedom in deciding how to pre-process the data and how to define "market segment." (Is it a group of correlated interests? A cluster? A principal component? Etc. You decide the answer to this question---don't ask me!) Just use the data to come up with some interesting, well-supported insights about the audience and give your client some insight as to how they might position their brand to maximally appeal to each market segment.

```{r}
social_marketing <- read.csv("~/Documents/Coding/eco395m-exercises/exercise_4/social_marketing.csv")


```

##  3. Association rules for grocery purchases